Contention Management for 5G

### 2017-04-27 article/article.tex
modifications dans la partie sans waiting times

	We consider a directed graph $G=(V,A)$ modelling a network. Each arc  $(u,v)$ in $A$ is labeled by an integer $dl(u,v) \geq 1$ that we call the delay and
	which represents the number of time slots taken by a signal to go from $u$ to $v$ using this arc. 
	%Note that for any arc $(u,v)$, $Dl(u,v)=Dl(v,u)$. Dominique voulait être plus général, on mettera cette propriété dans notre topologie

@@

\end{tikzpicture}
      \end{center}
       \caption{A routage graph with $(0,\dots,0)$ as a $(2,1)$-periodic affectation}
      \end{figure}

@@

	\begin{lemma} 
	 For any odd $P$, there is a routage graph such that there is $(2,1)$-periodic affectation but no $(P,1)$-periodic affectation.
	\end{lemma}
\begin{proof}

@@

      The message size $\tau$ in this process time, if we want to consider the time between the emission of the first slot and the reception of the last slot of the message. Since $\tau$ is the same for every routes, it does not change the calculations.

@@

We give two alternate proofs that PRA is $\NP$-complete.
The first one works for conflict depth $2$ and is minimal in this regards since we later prove that for conflict depth one,
it is easy to solve PRA. The second one reduces the problem to graph coloring and implies inapproximability when one tries to minimize the parameter $P$. \\

@@

  \subsection{Tractable cases of PRA}

    In this section, we present a few polynomial cases, in particular when the conflict depth is one.

    The problem PALL seems to be hard to solve, even for very simple classes of graphs, as seen in the next section.

@@

  In this subsection, we deal with a simpler version of the problem PALL.
  We ask for a $(P,\tau)$-periodic affectation {\bf with all waiting times equal to $0$} and we call this restricted problem  ??????????????? . 
  In that case $MT({\cal M})$ is equal to twice the delay of the longer route, thus $T_{max}$ is not relevant anymore. 
  Since $w_i=0$, choosing $m_i$, the offset of the route from $s_i$ to $t_i$, also sets the offset of the route from $t_i$ to $s_i$ to $m_i + \lambda(r_i) \mod P$.
  Remark that we can assume that the weight of the arcs from $s_i$ to $c_s$ are all equal to $0$, since we can change 
  all $m_i$ into $m_i - t(c_s,r_i) \mod P$ and obtain the same affectation.

@@

     By definition, there are no collision in the forward period and if the period is long enough, 
     it is easy to see that the order of the messages are the same as in the forward period and that no collision can occur.

      \begin{proposition}
       The algorithm Shortest-Longest solves ??????????? if $n\tau + 2(\lambda(r_{s_{n-1}}) - \lambda(r_{s_0})) \leq P$.
      \end{proposition}
      \begin{proof}
       Since $m_{s_i} = i\tau$, $[t(c_s,r_{s_i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.
       We assume that $\lambda(r_{s_0}) = 0$, by removing $\lambda(r_{s_0})$ to every route, it does not change their order
       or the result of this algorithm.

      We have  $[t(c_s,r_{t_i})] = \{2 \lambda(r_{s_i}) + i\tau, \dots,  2 \lambda(r_{s_i}) + (i+1)\tau -1\}$ since $2 \lambda(r_{s_{n-1 }}) + n\tau \leq P$.
       Since $ \lambda(r_{s_i}) \leq  \lambda(r_{s_{i+1}})$ by construction, we have  $2 \lambda(r_{s_i}) + (i+1)\tau -1 < 2 \lambda(r_{s_{i+1}}) + (i+1)\tau$
       which proves that there are no collision on the backward period. 
      \end{proof}

@@

    \begin{proposition}
    There is an algorithm, which finds a $(P,\tau)$-periodic affectation in time $O(n^2)$ when $P \geq 3n\tau$,
    where $n$ is the number of routes.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive macro slots of $\tau$ time slots. The algorithm works by chosing an offset for each route in the following way: try all offests which put the message in a yet not used macro slot in the forward
     period. An offset also fixes the position of the message in the backwward period, chose the first one which does not create a collision. We now prove that this algorithm always find an offset for all routes when $P \geq 3n\tau$.

@@

      \begin{center}
      \includegraphics[scale=0.3]{ex3nt.png}
      \end{center}
	\begin{algorithm}[H]
	\caption{Greedy affectation}
	\begin{algorithmic}
	\REQUIRE ${\cal R}_{\cal C}$, period $P$
	\ENSURE A P-periodic affectation in p $\leq P$, or FAILURE
	\STATE $P1[P/\tau]$ macro slots of size $\tau$ in forward period.
	\STATE $P2[P]$ slots backward period.
	\FORALL{source $s_i$ in S }

	\FORALL{macro slot j in P1}

	\IF{$P1[j]$ is free AND the corresponding slots in P2 are free}

	\STATE $m_{s_i} \leftarrow j.\tau$
	\ENDIF

	\IF{No intervals are found for $s_i$}
	\STATE return FAILURE
	\ENDIF
	\ENDFOR

	\ENDFOR

This algorithm, contrarily to the previous one, may work well, even when the condition $P \geq 3n\tau$ is not true.
In fact, experimental data in Subsection~\cite{} suggest that the algorithm works well as soon as $P \geq ????? \tau$.

@@

% 	    
      We now present an exhaustive search algorithm, which tries to set the offsets in all possible ways until it has found
      a $(P,\tau)$-periodic affectation. The algorithm is a depth-first traversal of a tree whose leaves are the $(P,\tau)$-periodic
      affectations or failures and the nodes are partial solutions. A partial solution $S$ is a choice of a starting time for a subset of the routes such that none of these routes collide in the forward or backward period. For simplicity we will also denote by $S$ the set of routes whose offsets are fixed by $S$. A partial solution $S'$ extends $S$ if $S' = S \cup \{r'\}$ and for all $r \in S$, $m_r + \tau \leq m_{r'}$, that is the route $r'$ comes after all routes in $S$ in the forward period. In the tree, the children of a partial solutions are all partial solutions which extend it. Therefore a node corresponding to $k$ fixed routes can have as many as $(n-k)P$ children. Since the tree is of depth $n$, we can have as many as $n!P^n$ elements and while $n$ is often small, $P$ may be large.  Therefore we have to find cuts in the tree to avoid to explore it entirely and henceforth make the algorithm practical. We propose the three following ones, the first two being useful when the period is small with regards to the number of routes:

      \todo{       Trouver un meilleur vocabulaire, peut-être avec des intervalles triés pour représenter les solutions partielles}

      \begin{enumerate}
       \item We maintain a counter $c_f$ which counts the number of routes which can be placed in the forward
       period. If at some point this number is less than the number of routes which are not in the partial solution,
       it is a failure and the algorithm backtracks. At the beginning of the algorithm $c_f = P / \tau$.
       Then when a route is fixed at some offset $m_r$, larger than all previous offsets, then $c_f = (P - m_r) / \tau $.

       \item We maintain a counter $c_b$ which counts the number of routes which can be placed in the backward
       period. If at some point this number is less than the number of routes which are not in the partial solution,
       it is a failure and the algorithm backtracks. At the beginning of the algorithm $c_f = P / \tau$.
       Consider a partial solution with $k$ routes and denote by $a_i$ their offset in the backward period.
       The $a_i$ are in increasing order and without loss of generality we assume that $a_1 = 0$. Remark that the number of routes which can be placed in the interval of free slots $\{a_i+\tau, a_{i+1}-1\}$
       is $(a_{i+1} -a_i) / \tau) -1 $. Therefore the total number of routes which can be placed in the backward period is $\displaystyle{\left( \sum_{i=1}^{k-1} (a_{i+1} -a_i)/\tau - 1 \right)} + (P - a_k) / \tau $.
       The counter $c_b$ can be maintained in constant time since each time a route is added, an interval is split into two and we can 
       use these two new intervals only to update its value.

%        In the forward windows, each message is placed after the previous one, that is $m_{s_i} > m_{s_j}$ (here, $s_i$ is the source placed before $s_j$ by the algorithm). Each time a new message is scheduled, there is some possible lost slots between $m_{s_i}+\tau$ and $ m_{s_j}$. We hold a counter of remaining slots and decrease it from the number of lost slots each time a new message is scheduled. Thus, if the number of the remaining slots in the forward period is lower than the number of slots needed to put the remaining messages, we can consider that the actual partial solution is not good.
%        \item  Each time we add a new message in the backward period, we cut a free interval in two new smaller free intervals, plus the scheduled message. Each free intervals has a number of macro slots (ex: 2,6). If the sum of all the integer parts of the macro slots is lower than the number of remaining messages to schedule, the actual partial solution is not good.
       \item 
       Let $S_1$ and $S_2$ be two partial solutions with the same subset of fixed routes, but different offsets.
       A sequence of extensions of a partial solution $S$ is a sequence of pair of route and offset, which define a path from $S$ in the tree. We say that $S_1$ dominates $S_2$, if any sequences of extensions of $S_2$, is a sequence of extensions of $S_1$.
       Then if there is a solution in $S_2$, there must be also one in $S_1$, therefore if we detect that some node is dominated in the tree we can backtrak when we encounter it.

       %Meilleure explication notion de slots utilisables. Si une solution partielle a strictement moins de slots utilisables (vu en tant qu'ensemble) qu'une autre avec les mêmes routes alors elle est dominée

       Consider a partial solution $S$ which we extend into $S' = S \cup \{ r \}$ such that the offset of 
       $r$ is the smallest possible, denoted by $m$.
       Say that the offset of $r$ in the backward period is $m+ \lambda$ and that the end of the message before $m + \lambda$ in the backward period is $a$. Then all extensions of $S$ into $S'' = S \cup \{ r \}$ with $ m < m_r < a + \tau - \lambda$.
       are dominated by $S'$. Indeed, $S''$ have striclty less empty slots in the forward period and in the backward period, there are only two intervals ...

       \todo{       Mettre un dessin !!!!  Dire comment la coupe 3 interragit bien avec la 1, car elle force $m_i$ à croitre vite, ce qui
       permet d'appliquer rapidement la coupe}

@@

    \subsubsection{Results}
      The following performance evaluation results compares the previous algorithms. Since we try to find a polynomial time algorithm for the problem PALL, Exhaustive Generation is not a good solution. However, Greedy $3.n.\tau$ and Shortest-Longest are some greedy heuristics and gives us a polynomial complexity. We tried to determine which algorithm gives a solution with the minimal period.
      In a first time, we will have a look at the results with the parameters corresponding to Cloud-RAN context.
      \paragraph{Short routes}
      To correspond with the C-RAN applications, we made some simulations on graphs in which $\lambda(c_s,r_{s_i})$ and $\lambda(c_t,r_{t_i})$, are drawn following an uniform act between 0 and 700 slots, that is, at the most, 5 km between the RRH and its BBU. The messages size $\tau$ is to 2500 slots, that is, approximately 1.228 Gbps of data for each route. The size of a slot is given by the time taken to send 64 bytes on the network, in which the links has 10Gbps throughput.

      To find the minimal period on a given instance with an algorithm, it is necessary to run the algorithm with some different periods and take the shortest one that does not fail. Since the existence of a $(P-\tau)-periodic affectation)$ is not monotone with regard to P (see section \ref{nonmonotone}), a dichotomous search is not appropriate. Thus, for each algorithm, we made a linear search of the minimal period for which, there is a solution on the given instance. The following figure shows us the average periods on 1000 graphs, for a growing number of routes.
      The results of the Exhaustive generation can be considered as a lower bound, indeed, for a given period, this algorithm finds the solution, if it exists.
      As an information, we drew the $\tau.n$ (i.e. the size taken by the messages in the period) and the $3.(\tau.n) $ lines.

@@

>>>>

	We consider a directed graph $G=(V,A)$ modeling a network. Each arc  $(u,v)$ in $A$ is labeled by an integer $dl(u,v) \geq 1$ that we call the delay and
	which represents the number of time slots taken by a signal to go from $u$ to $v$ using this arc. 
	%Note that for any arc $(u,v)$, $Dl(u,v)=Dl(v,u)$. Dominique voulait être plus général, on mettera cette propriété dans notre topologie

@@

\end{tikzpicture}
      \end{center}
       \caption{A routage graph with $(0,0,0)$ as a $(2,1)$-periodic affectation}
      \end{figure}

@@

	\begin{lemma} \label{lemma:monotonic}
	 For any odd $P$, there is a routage graph such that there is $(2,1)$-periodic affectation but no $(P,1)$-periodic affectation.
	\end{lemma}
\begin{proof}

@@

      In the process time, we count the time between the first time at which a message is emitted and 
      the first time at which the message comes back. Alternatively we could consider the time between the emission of the first slot and the reception of the last slot of the message, which would add $\tau$ to the process time.
      However, since all messages are of size $\tau$, it will not change the problem we solve in the rest of the article and that we now introduce.

@@

We give two alternate proofs that PRA is $\NP$-complete.
The first proof works already for conflict depth two. Remark that for conflict depth one,
the graph can be seen as a set of disjoint pair of routes, on which PRA and PALL can be solved in linear time. 
 The second proof reduces the problem to graph coloring and implies inapproximability when one tries to find the smallest possible $P$. \\

@@

%     
%   \subsection{Tractable cases of PRA}
%     
%     In this section, we present a few polynomial cases, in particular when the conflict depth is one.
%     When the conflict depth is one, every route can share an arc with only a single other route. 
%     Therefore 
%     
%     
%     The problem PALL seems to be hard to solve, even for very simple classes of graphs, as seen in the next section.

@@

  In this subsection, we deal with a simpler version of the problem PALL.
  We ask for a $(P,\tau)$-periodic affectation {\bf with all waiting times equal to $0$} and we call this restricted problem  
  {\bf Periodic Assignment for Zero Latency} or PAZL. 
  In that case $MT({\cal M})$ is equal to twice the delay of the longest route, thus $T_{max}$ is not relevant anymore. 
  Since $w_i=0$, choosing $m_i$, the offset of the route from $s_i$ to $t_i$, also sets the offset of the route from $t_i$ to $s_i$ to $m_i + \lambda(r_i) \mod P$.
 We assume from now on that the weight of the arcs from $s_i$ to $c_s$ are all equal to $0$, since we can change 
  all $m_i$ into $m_i - t(c_s,r_i) \mod P$ and obtain the same affectation.

@@

     By definition, there are no collision in the forward period and if the period is long enough, 
     it is easy to see that in the backward period the order of the messages are the same as in the forward period and that no collision can occur.

      \begin{proposition} When $n\tau + 2(\lambda(r_{s_{n-1}}) - \lambda(r_{s_0})) \leq P$, the algorithm Shortest-Longest solves PAZL.
      \end{proposition}
      \begin{proof}
       Since $m_{s_i} = i\tau$, $[t(c_s,r_{s_i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

       We assume that $\lambda(r_{s_0}) = 0$, by removing $\lambda(r_{s_0})$ to every route, it does not change the order on the routes
       nor the result of this algorithm.
      We have  $[t(c_s,r_{t_i})] = \{2 \lambda(r_{s_i}) + i\tau, \dots,  2 \lambda(r_{s_i}) + (i+1)\tau -1\}$ since $2 \lambda(r_{s_{n-1 }}) + n\tau \leq P$.
       Since $ \lambda(r_{s_i}) \leq  \lambda(r_{s_{i+1}})$ by construction, we have  $2 \lambda(r_{s_i}) + (i+1)\tau -1 < 2 \lambda(r_{s_{i+1}}) + (i+1)\tau$
       which proves that there are no collision on the backward period. 
      \end{proof}

@@

    \begin{proposition}
    Let  $n$ be the number of routes and let $ 3n\tau \leq P$, then there is an algorithm which finds a $(P,\tau)$-periodic affectation in time $O(n^2)$.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive macro slots of $\tau$ time slots. The algorithm works by choosing an offset for each route in the following way: try all offests which put the message in a yet not used macro slot in the forward
     period. The choice of an offset also fixes the position of the message in the backward period, chose the first one which does not create a collision. We now prove that this algorithm always find an offset for all routes when $P \geq 3n\tau$.

@@

      \begin{center}
      \includegraphics[scale=0.3]{ex3nt.png}
      \end{center}
% 	\begin{algorithm}[H]
% 	\caption{Greedy affectation}
% 	\begin{algorithmic}
% 	\REQUIRE ${\cal R}_{\cal C}$, period $P$
% 	\ENSURE A P-periodic affectation in p $\leq P$, or FAILURE
% 	\STATE $T$ a table of the macro slots of size $\tau$ in the forward period.
% 	\STATE $L$ a list of free intervals in the backward period%$P2[P]$ slots backward period.
% 	\FORALL{source $s$ in S}
% 
% 	\FORALL{free intervals $[a,b]$ in $L$}
% 	\FORALL{ $a/\tau - \lambda(s) <j< b/\tau - \lambda(s)$ }
% 	\IF{ $T[j] == FREE$}
% 	\STATE $m_{s} \leftarrow j.\tau$
% 	\STATE $T[j] = USED$
% 	\STATE update $[a,b]$ in $L$
% 	\STATE BREAK
% 	\ENDIF
% 	\ENDFOR
% 	\ENDFOR
% % 	
% % 	\IF{No intervals are found for $s_i$}
% % 	\STATE return FAILURE
% % 	\ENDIF
% % 	\ENDFOR
% 
% 	\ENDFOR
% 
% 	\end{algorithmic}
% 	\end{algorithm}
\todo{J'ai commencé à modifier l'algo pour qu'il colle, mais ça n'est pas encore ça. 
pas sur que ça vaille la peine de le mettre, je le commente pour l'instant.}

This algorithm, contrarily to the previous one, may work well, even when the condition $P \geq 3n\tau$ is not true.
In fact, experimental data in Subsection~\ref{sec:experimental_results} suggest that the algorithm works well as soon as $P \geq 1.5 \tau$ for small routes.

\todo{vérifier cette affirmation en faisant une expérience et une régression linéaire}

@@

% 	    
      We now present an exhaustive search algorithm, which tries to set the offsets in all possible ways until it has found
      a $(P,\tau)$-periodic affectation. We have $n$ routes denoted by $\{0,\dots,n-1\}$. A partial solution $S$ is 
      a partial function from $[n]$ to $\{0,1,\dots,P-\tau -1\}$ which sets a starting time for a subset of the routes $R(S) \subseteq [n]$,
      such that there are no collisions for these routes.  A partial solution $S'$ extends $S$, if it sets one more routes than $S$ to a larger starting time than all routes in $S$, that is $R(S') = R(S) \cup \{r'\}$ and for all  $r \in R(S)$, $S(r) + \tau \leq S'{r'}$. We define a tree whose nodes are partial solutions and such that the root is the empty partial solution and 
      the children of a partial solution are the partial solutions which extend it. All solutions to our problem will be 
      leaves in this tree, therefore our algorithm is simply a depth-first search of the tree.

      Remark that a node $S$ with $|R(S)| = k$ can have as many as $(n-k)P$ children. Since the tree is of depth $n$, the tree may have as many as $n!P^n$ elements and while $n$ is often small, $P$ may be large which makes its traversal intractable.  Therefore we have to find cuts in the tree to avoid to explore it entirely and henceforth make the algorithm practical. Cuts correspond to the detection of subtrees which contain no solutions and can thus be skipped.   
      We now propose three cuts, the first two being particularly useful when the network is loaded ($n\tau$ is not much smaller than $P$).

      \begin{enumerate}
       \item We consider the number of slots which can be used by routes not yet fixed by a partial solution in the 
       \emph{forward period}. When we extend a solution into $S$ by a route at offset $m$, then at most $(P - m) / \tau $ routes can still be set without collisions in the forward period. If that value is less than the number of routes which are not in $R(S)$, it is a failure and the algorithm backtracks.

       \item 
       For the next two cuts, we need to define the notion of the useful slots of a partial solution $S$ in the \emph{backward period}: a slot is said to be useful, if it is not used by a message set by $S$ in the backward period and it belongs to an interval of at least $\tau$ such slots. Useful slots are positions which can be used when extending $S$. We will denote by $([a_i,b_i[)_{i\leq l}$ the ordered sequence of intervals of useful slots of $S$. Without loss of generality we can assume that all $a_i, b_i \leq l$. The number of message of size $\tau$ which can be set in the useful slots of $S$ is thus  $\displaystyle{ \sum_{i=0}^{l} (b_{i} -a_i)/\tau } $. If that value is less than the number of routes which are not in $R(S)$, it is a failure and the algorithm backtracks. Notice that the list of intervals of useful slots and the value $\displaystyle{ \sum_{i=0}^{l} (b_{i} -a_i)/\tau } $ can be maintained in constant time, since each time a route is added, we only need to split an interval of useful slots into at most two such intervals.

       \item 
       Let $S_1$ and $S_2$ be two partial solutions with $R(S_1) = R(S_2)$. Let $US_1$ (respectively $US_2$) be the set of useful slots of $S_1$ (resp. $S_2$). We say that $S_1$ dominates $S_2$ if the largest offset fixed in $S_1$ is smaller than the one in $S_2$ and $US_2 \subseteq US_1$. Remark that any valid sequence of extension of $S_2$ (fixing offsets of routes in the complementary of 
       $R(S_2)$) is also a valid sequence of extensions of $S_1$. Therefore if the tree rooted at $S_2$ contain a solution, then $S_1$ contains one too. Hence, in our exhaustive search of the tree of partial solutions, we can skip the tree rooted at $S_2$.

       We now explain how we can detect some partial solutions which are dominated so that we do not explore their subtrees.
       Consider a partial solution $S$ which we extend into $S'$ by setting the offset of the route $r$ to be the smallest possible. The offset of $r$ in the backward period is $S'(R)+ \lambda$ and we denote the end of the message before before by $a$. Then all extensions of $S$ into $S''$ with $S'(r)  < S''(r) < a + \tau - \lambda$ are dominated by $S'$. Therefore when computing the extension of $S$, we first build $S'$ and then $S''$ with $S''(r) =  a + \tau - \lambda$ , skipping all values in between.

@@

      The third cut works well in conjunction with the first one since it makes the offsets grow quickly and 
      which makes the first cut more likely to apply. A last cut could be implemented, by computing whether for every 
      route not in $R(S)$ the set of possible positions in the backward period and checking whether one s contained in the useful slots of $S$.

    \subsubsection{Experimental Results}\label{sec:experimental_results}

      The following performance evaluation results compares the three presented algorithms.
      Notice that both the Greedy algorithm and the Shortest-Longest are polynomial time algorithm but are not always able to
      find optimal solutions. On the other hand, the exhaustive search will find it but it works in exponential time.
      We will compare the performance of the algorithms in two different regimes: the routes are either smaller than 
      the size of a message or they are arbitrary long.

            The graphs 
      To correspond with the C-RAN applications, we made some simulations on graphs in which $\lambda(c_s,r_{s_i})$ and $\lambda(c_t,r_{t_i})$, are drawn following an uniform act between 0 and 700 slots, that is, at the most, 5 km between the RRH and its BBU. The messages size $\tau$ is to 2500 slots, that is, approximately 1.228 Gbps of data for each route. The size of a slot is given by the time taken to send 64 bytes on the network, in which the links has 10Gbps throughput.

      \todo{Mettre les données concrètes plus proprement et éventuellement ne pas les mettre si on envoie pas à un truc de télécom}

      \paragraph{Short routes}

       We chose parameters coming from our C-RAN settings. A message is of $2500$ slots, while the values $\lambda(c_t,r_{t_i})$ are drawn uniformly between $0$ and $700$ slots. For each of the three different algorithms and for a given instance we look for the minimal period $P$ for which the problem has a $(P,\tau)$ periodic affectation. We then compute the average of these minimal periods for $1000$ random star topologies for different number of routes. To find the minimum period, we need to do a linear search on the value $P$, a dichotomy would not work because of Lemma~\ref{lemma:monotonic}.

%our network has a throughput of 10Gbps throughput and 
%       
%       
%       The messages size $\tau$ is of 2500 slots, that is, approximately 1.228 Gbps of data for each route. The size of a slot is given by the time taken to send 64 bytes on the network, in which the links has 10Gbps throughput.

      Since in this setting the exhaustive generation was always finishing in less than ten seconds, the results corresponding to this
      algorithms are optimal and are a lower bounds for the period found for the other algorithms.
      As an information, we draw the $f(n) = n\tau$, which is an absolute lower bound on the period and $g(n) = 3n\tau$ lines which is an upper bound on the period found by the greedy algorithm.

@@

      \paragraph{Long routes}
